An essentially algebraic formulation of dependent type theory

### 2015-01-19 Rules/rules_extension.tex
fixed broken forward reference to term extension

In \autoref{extension-on-terms} we will also define an extension operation
on terms, which defines a term $\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for
every $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$. This
operation is defined by using the substitution operation and identity terms.

>>>>

Using identity terms and substitution, it is also possible to define an 
extension operation on terms, which defines a term 
$\jterm{\Gamma}{{A}{P}}{\tmext{x}{u}}$ for every $\jterm{\Gamma}{A}{x}$ and 
$\jterm{\Gamma}{\subst{x}{P}}{u}$. Thus, there is a formal definition of the
\emph{pair} $\tmext{x}{u}$. We shall not go into this, in the current paper,
but it will be very useful to have when one further develops the theory of
E-systems.

### 2015-01-19 main.tex
fixed broken forward reference to term extension

\includeonly{Algebras/article-models-cat}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Internal models of type theory and internal higher categories for the
Univalent Foundations}
\author{Egbert Rijke}
\date{\today}

@@

\tableofcontents

\input{Introduction/article-models-intro}

\part{Type theory}

In this part we develop type theory from the ground up. We start with a type
theory without any of the basic constructors. This is the theory of contexts
families and terms which has the basic operations of extension, weakening,
substitution and identity terms. Type theory before type constructors has not
been studied very much. Dependent product types or even universes tend to make
an early appearance in just about any presentation of type theory.
A noteable exception is the theory of categories with families of Dybjer
in \cite{Dybjer1996}, which has been elaborated on further by Dybjer and
Clairambault in \cite{DybjerClairambault2011} and in unpublished work by
Awodey \cite{Awodey2013} on natural models of type theory,
which makes the connection between categories with families and representable
transformations of presheaves. In the way type theory is presented by the
Univalent Foundations Project in \cite{TheBook}, which seems to have won the race of introducing universes
as early as possible hands down, it seems entirely unfeasible to study type
theory without type constructors. Also, the 
proof-assistant \Coq\ {\color{red}(and \Agda too?)} has universes and dependent product and pair types 
built-in, making it impossible to study type theory without type constructors in 
that environment.

@@

After we have described the E-system, which is the flavor of type theory without
type constructors presented in \autoref{tt}, we will demonstrate in \autoref{ttderived}
that the theory gives rise to a rich categorical structure. The notions and
properties we derive here will be essential for the further work on internal
models for E-systems, presented in \autoref{part:models}.

\include{Rules/rules}

\include{Derivations/article-models-ttderived}

\part{Internal Models}\label{part:models}

\include{Internal_models/article-models-ttinternal}

\part{Type constructors}

\include{Dependent_function_types/article-models-ttdp}

\include{Universes/article-models-ttuniverses}

\include{Junk/article-models-ttind}

\part{Categorical semantics}

\include{Algebras/article-models-cat}

>>>>

%\includeonly{Algebras/article-models-cat}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{An essentially algebraic formulation of dependent type theory}
\author{Egbert Rijke}
\date{\today}

@@

%\tableofcontents

%\input{Introduction/article-models-intro}

%\part{Type theory}

\begin{comment}
In this part we develop type theory from the ground up. We start with a type
theory without any of the basic constructors. This is the theory of contexts
families and terms which has the basic operations of extension, weakening,
substitution and identity terms. Type theory before type constructors has not
been studied very much. Dependent product types or even universes tend to make
an early appearance in just about any presentation of type theory.
A noteable exception is the theory of categories with families of Dybjer
in \cite{Dybjer1996}, which has been elaborated on further by Dybjer and
Clairambault in \cite{DybjerClairambault2011} and in unpublished work by
Awodey \cite{Awodey2013} on natural models of type theory,
which makes the connection between categories with families and representable
transformations of presheaves. In the way type theory is presented by the
Univalent Foundations Project in \cite{TheBook}, which seems to have won the race of introducing universes
as early as possible hands down, it seems entirely unfeasible to study type
theory without type constructors. Also, the 
proof-assistant \Coq\ {\color{red}(and \Agda too?)} has universes and dependent product and pair types 
built-in, making it impossible to study type theory without type constructors in 
that environment.

@@

After we have described the E-system, which is the flavor of type theory without
type constructors presented in \autoref{tt}, we will demonstrate in \autoref{ttderived}
that the theory gives rise to a rich categorical structure. The notions and
properties we derive here will be essential for the further work on internal
models for E-systems, presented in \autoref{part:models}.
\end{comment}

\input{Rules/rules}

%\include{Derivations/article-models-ttderived}

%\part{Internal Models}\label{part:models}

%\include{Internal_models/article-models-ttinternal}

%\part{Type constructors}

%\include{Dependent_function_types/article-models-ttdp}

%\include{Universes/article-models-ttuniverses}

%\include{Junk/article-models-ttind}

%\part{Categorical semantics}

\input{Algebras/article-models-cat}
